---
title: "Nonlinearities in fix expo windows"
author: "Yanelli Nunez"
date: "9/30/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

*Objectives* 

- Used a natural spline with 3 degrees of freedom to test for non-linearities in the outcome-exposure relationship. 

- Use the AIC to compare model fitness between the model with the natural spline and the model without it

*Results*

For all averages (1, 5, and 10yrs), the conditional logistic linear models have lower AIC which means that models with the linear exposure have better goodness of fit based on AIC


## i.  Loading needed packages

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))

# load functions
 source(paste0(code.folder, 'models/functions_yanelli.R'))

# load data
 source(paste0(code.folder, 'models/m_00_load_datasets.R'))
```

## ii. View data summaries

```{r}
summary(data_1yr)


summary(data_5yrs)


summary(data_10yrs)

library(AICcmodavg)

```

## 1. Check nonlinearities in the conditional logistic models



### 1a. 1yr exposure average

Model with linear term has a lower AIC 

Site with info about using AIC to compare models:

https://www.scribbr.com/statistics/akaike-information-criterion/#:~:text=The%20AIC%20function%20is%202K,it%20is%20being%20compared%20to.
```{r}

# linear

clogit.1yr <- survival::clogit(ck ~ PM25_1yr + 
                           family_SES + 
                           place_birth +
                           civst +
                           residence +
                              parish_ses +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_1yr) 


summary(clogit.1yr)

# nonlinear

clogit.1yr.ns <- survival::clogit(ck ~ ns(PM25_1yr, df = 3) + 
                           family_SES + 
                           place_birth +
                           civst +
                           residence +
                             parish_ses +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_1yr) 


summary(clogit.1yr.ns)



# AIC - lower AIcC means  higher model fitness
aictab(cand.set = list(clogit.1yr, clogit.1yr.ns), modnames = c('linear', 'nonlinear'))

```

### 1b. 5 years exposure average

Model with linear exposure term has better AIC
```{r}

# linear
clogit.5yr <- clogit(ck ~ PM25_5yr + 
                           family_SES + 
                           place_birth +
                           civst +
                       residence +
                        parish_ses +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_5yrs) 
summary(clogit.5yr)


# nonlinear
clogit.5yr.ns <- clogit(ck ~ ns(PM25_5yr, df = 3) + 
                           family_SES + 
                           place_birth +
                           civst +
                       residence +
                          parish_ses +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_5yrs) 
summary(clogit.5yr.ns)

# Get AICs for each model

aictab(cand.set = list(clogit.5yr, clogit.5yr.ns), modnames = c('linear', 'nonlinear'))

```

### 1c. 10 years exposure
In this case the AIC values are very similar but the linear model still has a better AIC
```{r}

# Linear
clogit.10yrs <- clogit(ck ~ PM25_10yr + 
                           family_SES + 
                           place_birth +
                           civst +
                         residence +
                          parish_ses +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_10yrs) 

summary(clogit.10yrs)

# Nonlinear

clogit.10yrs.ns <- clogit(ck ~ ns(PM25_10yr, df =3) + 
                           family_SES + 
                           place_birth +
                           civst +
                         residence +
                            parish_ses +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_10yrs) 

summary(clogit.10yrs.ns)

# AIC comparison
aictab(cand.set = list(clogit.10yrs, clogit.10yrs.ns), modnames = c('linear', 'nonlinear'))

```

