---
title: "y_05_pm2.5_lag_analysis"
author: "Yanelli Nunez"
date: "10/4/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:

- Ran distributed  lag models and two different lag constrains (3 and 4 degrees of freedom). The exposure-response function  will be set as linear because the models for the averaged exposures indicate no deviation from linearity in the dose-response. 

- To  determined df on the lag constrain I used AIC to chose the model with the best fitness

- I ran the model with the best AIC twice--with and without parish SES adjustment 

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))

# load functions
 source(paste0(code.folder, 'models/functions_yanelli.R'))

# load data
 source(paste0(code.folder, 'models/m_00_load_datasets.R'))

```

## 1. View data


```{r}

summary(data_lags)

```

## 2. Organize data 

- Arrange data into the proper format for the lag analysis
- 768 individuals with parish-level SES missing were removed

```{r}

lag_wide <- data_lags %>%
 group_by(nr, lobnr) %>%
  mutate(lags = as.character(lags), 
         lags = paste0("lag_", lags)) %>%
  ungroup() %>%
 spread(lags, mean_pm2.5)

summary(lag_wide)

head(lag_wide, 8)

# dataset excluding cases with missing parish-level SES
excluded_neigh_SES <- lag_wide %>%
  drop_na(parish_ses)
  
```

 
## 3. Create matrix of exposure histories 

For lags 1-10

```{r exposure histories}

# matrix with 10 lags
exp_matrix <- lag_wide %>%
  dplyr::select(33:43) %>%
  dplyr::select(lag_1, lag_2, lag_3, lag_4, lag_5, lag_6, lag_7, lag_8, lag_9, lag_10)

head(exp_matrix, 8)

exp_excluded_neigh_SES <- excluded_neigh_SES %>%
  dplyr::select(33:43) %>%
  dplyr::select(lag_1, lag_2, lag_3, lag_4, lag_5, lag_6, lag_7, lag_8, lag_9, lag_10)


```


## 4. Prep matrixes for models

### 4a. Crossbasis

All crossbases have a linear dose-response relationship

I tested two different constrains on the lag: 

 - Lag constrained with 4 df
 - Lag constrained with 3 df
 

```{r}

# Crossbasis with 10 lags 

# 4 df 
cb.lin.4 <- crossbasis(exp_matrix, 
                 lag=c(1,10),
                 argvar=list(fun = "lin"), # form of the dose-response within each lag
                 arglag=list(fun= "ns", df = 4 )) # the form of the curve across all the lags, that is, the lag constrain 




# 3 df 
cb.lin.3 <- crossbasis(exp_matrix, 
                 lag=c(1,10),
                 argvar=list(fun = "lin"), # form of the dose-response within each lag
                 arglag=list(fun= "ns", df = 3 )) # the form of the curve across all the lags, that is, the lag constrain 


# 3 df excluding people with missing parish-level SES
cb.lin.3.excluded.SES <- crossbasis(exp_excluded_neigh_SES, 
                 lag=c(1,10),
                 argvar=list(fun = "lin"), # form of the dose-response within each lag
                 arglag=list(fun= "ns", df = 3 )) # the form of the curve across all the lags, that is, the lag constrain 



```


## 5. Lag Models

### 5a. Linear conditional logistic models with 10 lags 


#### i. Lag constrain with 4 df: 


```{r}

# run model
clogit.lin.4 <- survival::clogit(ck ~ cb.lin.4 + 
                               family_SES +
                               place_birth +
                               civst +
                               residence + 
                               strata(nr), method = "efron", data = lag_wide)

summary(clogit.lin.4)


# Predict
pred.lin.4 <- crosspred(
  cb.lin.4, # crossbasis
  clogit.lin.4, # model
  at = 1:30, # range values of PM2.5 to use for prediction select based on min/max pm2.5 concentrations in dataset
  bylag = 1, # lag increments 
  cumul = TRUE, # estimate cumulative effect
  cen = 0) # references for estimate


## Visuals

# 3-D plot
plot(pred.lin.4, zlab="OR", xlab="Exposure", ylab="Lag (years)",
       theta=40, phi=30, lphi=30)

# Plot lag exposure-response at exposure of 10ug/m3
plot(pred.lin.4, 
     var = 10, 
     col = "red", 
     ci.arg = list(density = 15, lwd =2),
      ylab = "OR at 10ug/m3 PM2.5", xlab = "lag (years)")



## Extract estimates per lag 
# I created the OR_lag() and OR_cumm() function to extract the per lag and cummulative estimates and organize them into a dataset to use later for figures. The functions aare in  the functions_yanelli.R  file  in this folder
OR_4df <- OR_lag(pred.lin.4, "4-df")


## Extract cumulative estimates for 1, 5 and 10 lags

OR_cumm_4df <- OR_cumm(pred.lin.4, "4-df")

```

#### ii. Lag constrain with 3 df: 

- This model has a lower AIC relative to the model with a lag constrain of 4df. So we select this models as the model for the main analysis

- This model is not adjusted for neighborhood-level SES and includes participants with missing neighborhood-level SES

```{r}

clogit.lin.3 <- survival::clogit(ck ~ cb.lin.3 + 
                               family_SES +
                               place_birth +
                               civst +
                               residence +
                               strata(nr), method = "efron", data = lag_wide)

summary(clogit.lin.3)



# Predict

pred.lin.3 <- crosspred(
  cb.lin.3, # crossbasis
  clogit.lin.3, # model
  at = 1:30, # range values of PM2.5 to use for prediction
  bylag = 1, # lag increments 
  cumul = TRUE, # estimate cumulative effect at each lag
  cen = 0) # references for nonlinear associations


# Visuals

# 3-D plot
plot(pred.lin.3, zlab="OR", xlab="Exposure", ylab="Lag (years)",
       theta=40, phi=30, lphi=30)

# Plot lag exposure-response at exposure of 10ug/m3

plot(pred.lin.3, 
     var = 10, 
     col = "red", 
     ci.arg = list(density = 15, lwd =2),
      ylab = "OR at 10ug/m3 PM2.5", xlab = "lag (years)")

## Extract estimates per lag

OR_3df <- OR_lag(pred.lin.3, "3-df")

## Extract cumulative estimates at each lag

OR_cumm_3df <- OR_cumm(pred.lin.3, "3-df")

```

### 5b. Compare models' AIC values 

- Compare the AIC values for the 3-df and 4-df lag constraint models
- The model with the 3df constrain has a better fit based on AIC

```{r}
aictab(cand.set = list(clogit.lin.3, clogit.lin.4), modnames = c('3df-constraint', '4df-constraint'))
```



### 5c. Lag constrain with 3 df adjusted for parish SES

- The model with a lag constrain with 3df has the lowest AIC so here I'm running the model again but adjusted for parish SES

- Because I'm adjusting for parish-level SES, this model does not include individuals without parish-level SES info

```{r}

clogit.lin.ses <- survival::clogit(ck ~ cb.lin.3 + 
                               family_SES +
                               place_birth +
                               civst +
                               residence +
                               parish_ses +
                               strata(nr), method = "efron", data = lag_wide)

summary(clogit.lin.ses)


# Predict

pred.lin.ses <- crosspred(
  cb.lin.3, # crossbasis
  clogit.lin.ses, # model
  at = 1:30, # range values of PM2.5 to use for prediction
  bylag = 1, # lag increments 
  cumul = TRUE, # estimate cumulative effect at each lag
  cen = 0) # references for nonlinear associations


# Visuals

# 3-D plot
plot(pred.lin.ses, zlab="OR", xlab="Exposure", ylab="Lag (years)",
       theta=40, phi=30, lphi=30)

# Plot lag exposure-response at exposure of 10ug/m3

plot(pred.lin.ses, 
     var = 10, 
     col = "red", 
     ci.arg = list(density = 15, lwd =2),
      ylab = "OR at 10ug/m3 PM2.5", xlab = "lag (years)")

## Extract estimates per lag

OR_3df_ses <- OR_lag(pred.lin.ses, "3-df w/parish ses")


## Extract cumulative estimates at each lag

OR_cumm_3df_ses <- OR_cumm(pred.lin.ses, "3-df with/parish ses")

```

### 5c. Lag constrain with 3 df without adjusting for parish SES

- This model doesn't adjust for parish-level SES AND EXCLUDES individuals with missing parish-level SES

```{r}

clogit.lin.excluded <- survival::clogit(ck ~ cb.lin.3.excluded.SES + 
                               family_SES +
                               place_birth +
                               civst +
                               residence +
                               strata(nr), method = "efron", data = excluded_neigh_SES)

summary(clogit.lin.excluded)


# Predict

pred.lin.exclud <- crosspred(
  cb.lin.3.excluded.SES, # crossbasis
  clogit.lin.excluded, # model
  at = 1:30, # range values of PM2.5 to use for prediction
  bylag = 1, # lag increments 
  cumul = TRUE, # estimate cumulative effect at each lag
  cen = 0) # references for nonlinear associations


# Visuals

# 3-D plot
plot(pred.lin.exclud, zlab="OR", xlab="Exposure", ylab="Lag (years)",
       theta=40, phi=30, lphi=30)

# Plot lag exposure-response at exposure of 10ug/m3

plot(pred.lin.exclud, 
     var = 10, 
     col = "red", 
     ci.arg = list(density = 15, lwd =2),
      ylab = "OR at 10ug/m3 PM2.5", xlab = "lag (years)")

## Extract estimates per lag

OR_3df_exclud <- OR_lag(pred.lin.exclud, "3-df w/out parish ses/indiv")


## Extract cumulative estimates at each lag

OR_cumm_3df_exclud <- OR_cumm(pred.lin.exclud, "3-df w/out parish ses/indiv")

```




## 7. Compile and save effect estimates

- I combined the effect estimates from the three models (4df and 3df lag constrain, and 3df lag constrained adjusted for parish SES ) into a single dataset both for the single lag effect estimates and the cumulative effect estimates. I saved the dataset as a .csv into file in the output folder

```{r}

# by lag effect estimates

per_lag_effects <- rbind(OR_3df, OR_4df) %>%
  rbind(., OR_3df_ses) %>%
  rbind(., OR_3df_exclud)

knitr::kable(per_lag_effects)

 write_csv(per_lag_effects, paste0(output.folder, "distributed_lags_OR.csv"))

# cumulative estimates 

cumm_effects <- rbind(OR_cumm_3df, OR_cumm_4df) %>%
  rbind(., OR_cumm_3df_ses) %>%
  rbind(., OR_cumm_3df_exclud)

knitr::kable(cumm_effects)

 write_csv(cumm_effects, paste0(output.folder, "cumm_lags_OR.csv"))


```

