---
title: "m_04_sex_stratified_analysis"
author: "Yanelli Nunez"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## i.  Loading needed packages

* Objective * 

Construct models stratified by sex (one model per sex). 

Male = 1
Female = 2



```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))

# load functions
 source(paste0(code.folder, 'models/functions_yanelli.R'))

# load data
 source(paste0(code.folder, 'models/m_00_load_datasets.R'))

```

## 1. Organize data 

- Arrange data into the proper format for the lag analysis


```{r}

lag_wide <- data_lags %>%
 group_by(nr, lobnr) %>%
  mutate(lags = as.character(lags), 
         lags = paste0("lag_", lags)) %>%
  ungroup() %>%
 spread(lags, mean_pm2.5)

summary(lag_wide)

head(lag_wide, 8)

# sex stratified datasets 
lag_male <- lag_wide %>%
  dplyr::filter(sex == 1)

lag_female <- lag_wide %>%
  dplyr::filter(sex ==2)

```

 
## 2. Create matrix of exposure histories 

For lags 1-10

```{r exposure histories}

# matrix with 10 lags

# male
m_exp_matrix <- lag_male %>%
  dplyr::select(33:43) %>%
  dplyr::select(lag_1, lag_2, lag_3, lag_4, lag_5, lag_6, lag_7, lag_8, lag_9, lag_10)



# female
f_exp_matrix <- lag_female %>%
  dplyr::select(33:43) %>%
  dplyr::select(lag_1, lag_2, lag_3, lag_4, lag_5, lag_6, lag_7, lag_8, lag_9, lag_10)


```

## 3. Crossbasis


All crossbases have a linear dose-response relationship

In the main analysis we found that the model with a lag constrained with 3df has a better fit. So here we use 3df for the sex stratified models


```{r}

# Crossbasis with 10 lags 

# male 
m.cb.lin <- crossbasis(m_exp_matrix, 
                 lag=c(1,10),
                 argvar=list(fun = "lin"), # form of the dose-response within each lag
                 arglag=list(fun= "ns", df = 3)) # the form of the curve across all the lags, that is, the lag constrain 


# female
f.cb.lin <- crossbasis(f_exp_matrix, 
                 lag=c(1,10),
                 argvar=list(fun = "lin"), # form of the dose-response within each lag
                 arglag=list(fun= "ns", df = 3)) # the form of the curve across all the lags, that is, the lag constrain 



```


## 4. Lag Models

Linear conditional logistic models with 10 lags and 3 df in the lag constrain


### 4a. Male 


```{r}

# run model
clogit.lin.male <- survival::clogit(ck ~ m.cb.lin + 
                               family_SES +
                               place_birth +
                               civst +
                               residence + 
                               strata(nr), method = "efron", data = lag_male)

summary(clogit.lin.male)


# Predict
pred.lin.male <- crosspred(
  m.cb.lin, # crossbasis
  clogit.lin.male, # model
  at = 1:30, # range values of PM2.5 to use for prediction select based on min/max pm2.5 concentrations in dataset
  bylag = 1, # lag increments 
  cumul = TRUE, # estimate cumulative effect
  cen = 0) # references for estimate


## Visuals

# 3-D plot
plot(pred.lin.male, zlab="OR", xlab="Exposure", ylab="Lag (years)",
       theta=40, phi=30, lphi=30)

# Plot lag exposure-response at exposure of 10ug/m3
plot(pred.lin.male, 
     var = 10, 
     col = "red", 
     ci.arg = list(density = 15, lwd =2),
      ylab = "OR at 10ug/m3 PM2.5", xlab = "lag (years)")



## Extract estimates per lag 
# I created the OR_lag() and OR_cumm() function to extract the per lag and cummulative estimates and organize them into a dataset to use later for figures. The functions are in  the functions_yanelli.R  file  in this folder
OR_lag_male <- OR_lag(pred.lin.male, "male")


## Extract cumulative estimates for 1, 5 and 10 lags

OR_cumm_male <- OR_cumm(pred.lin.male, "male")

```

### 4b. Female 

```{r}

# run model
clogit.lin.female <- survival::clogit(ck ~ f.cb.lin + 
                               family_SES +
                               place_birth +
                               civst +
                               residence + 
                               strata(nr), method = "efron", data = lag_female)

summary(clogit.lin.female)


# Predict
pred.lin.female <- crosspred(
  f.cb.lin, # crossbasis
  clogit.lin.female, # model
  at = 1:30, # range values of PM2.5 to use for prediction select based on min/max pm2.5 concentrations in dataset
  bylag = 1, # lag increments 
  cumul = TRUE, # estimate cumulative effect
  cen = 0) # references for estimate


## Visuals

# 3-D plot
plot(pred.lin.female, zlab="OR", xlab="Exposure", ylab="Lag (years)",
       theta=40, phi=30, lphi=30)

# Plot lag exposure-response at exposure of 10ug/m3
plot(pred.lin.female, 
     var = 10, 
     col = "red", 
     ci.arg = list(density = 15, lwd =2),
      ylab = "OR at 10ug/m3 PM2.5", xlab = "lag (years)")



## Extract estimates per lag 
# I created the OR_lag() and OR_cumm() function to extract the per lag and cummulative estimates and organize them into a dataset to use later for figures. The functions are in  the functions_yanelli.R  file  in this folder
OR_lag_female <- OR_lag(pred.lin.female, "female")


## Extract cumulative estimates for 1, 5 and 10 lags

OR_cumm_female <- OR_cumm(pred.lin.female, "female")

```

## 5. Compile and save effect estimates

- I combined the effect estimates from the two models (female and male) into a single dataset both for the single lag effect estimates and the cumulative effect estimates. I saved the dataset as a .csv into the output folder

```{r}

# by lag effect estimates

per_lag_effects <- rbind(OR_lag_female, OR_lag_male) %>%
dplyr::rename(model = lag_constr)

knitr::kable(per_lag_effects)

 write_csv(per_lag_effects, paste0(output.folder, "sex_distributed_lags_OR.csv"))

# cumulative estimates 

cumm_effects <- rbind(OR_cumm_male, OR_cumm_female) %>%
 dplyr::rename(model = lag_constr)

knitr::kable(cumm_effects)

 write_csv(cumm_effects, paste0(output.folder, "sex_cumm_lags_OR.csv"))

```

