---
title: "m_06_inpatient_only_lag_analysis"
author: "Yanelli Nunez"
date: "6/14/2022"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## i.  Loading needed packages

* Objective * 

Construct model that includes only inpatient hospitalizations

I used the variable "pattype" to select only inpatient hospitalizations

0 = "Never hospital patient"
1 = "inpatient"
2 = "outpatient"
3 = "Emergency


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))

# load functions
 source(paste0(code.folder, 'models/functions_yanelli.R'))

# load data
 source(paste0(code.folder, 'models/m_00_load_datasets.R'))

```

## 1. Organize data 


- Arrange data into the proper format for the lag analysis

Notes: 

Most cases were outpatient. See the breakdowns below:

inpatient = 1526
Outpatient = 2237
emergency = 11

```{r}

lag_wide <- data_lags %>%
 group_by(nr, lobnr) %>%
  mutate(lags = as.character(lags), 
         lags = paste0("lag_", lags)) %>%
  ungroup() %>%
 spread(lags, mean_pm2.5)

# check how many inpatients we have
lag_wide %>%
  dplyr::filter(ck == 1) %>%
  mutate(pattype = as.factor(pattype)) %>%
  summary()

head(lag_wide, 8)

# select only inpatient cases 
lag_inpatient <- lag_wide %>%
  dplyr::filter(pattype == 1) 

```

 
## 2. Create matrix of exposure histories 

For lags 1-10

```{r exposure histories}

# matrix with 10 lags

exp_matrix <- lag_inpatient %>%
  dplyr::select(33:43) %>%
  dplyr::select(lag_1, lag_2, lag_3, lag_4, lag_5, lag_6, lag_7, lag_8, lag_9, lag_10)


```

## 3. Crossbasis


All crossbases have a linear dose-response relationship

In the main analysis we found that the model with a lag constrained with 3df has a better fit. So here we use 3df for the age stratified models


```{r}

# Crossbasis with 10 lags 


cb.lin.inpatient <- crossbasis(exp_matrix, 
                 lag=c(1,10),
                 argvar=list(fun = "lin"), # form of the dose-response within each lag
                 arglag=list(fun= "ns", df = 3)) # the form of the curve across all the lags, that is, the lag constrain 



```


## 4. Lag Models

Linear conditional logistic models with 10 lags and 3 df in the lag constrain

```{r}

# run model
clogit.lin <- survival::clogit(ck ~ cb.lin.inpatient + 
                               family_SES +
                               place_birth +
                               civst +
                               residence + 
                               strata(nr), method = "efron", data = lag_inpatient)

summary(clogit.lin)


# Predict
pred.lin <- crosspred(
  cb.lin.inpatient, # crossbasis
  clogit.lin, # model
  at = 1:30, # range values of PM2.5 to use for prediction select based on min/max pm2.5 concentrations in dataset
  bylag = 1, # lag increments 
  cumul = TRUE, # estimate cumulative effect
  cen = 0) # references for estimate


## Visuals

# 3-D plot
plot(pred.lin, zlab="OR", xlab="Exposure", ylab="Lag (years)",
       theta=40, phi=30, lphi=30)

# Plot lag exposure-response at exposure of 10ug/m3
plot(pred.lin, 
     var = 10, 
     col = "red", 
     ci.arg = list(density = 15, lwd =2),
      ylab = "OR at 10ug/m3 PM2.5", xlab = "lag (years)")



## Extract estimates per lag 
# I created the OR_lag() and OR_cumm() function to extract the per lag and cummulative estimates and organize them into a dataset to use later for figures. The functions are in  the functions_yanelli.R  file  in this folder
OR_lag_inpat <- OR_lag(pred.lin, "inpatient")

knitr::kable(OR_lag_inpat)

 write_csv(OR_lag_inpat, paste0(output.folder, "inpatients_distributed_lags_OR.csv"))



## Extract cumulative estimates for 1, 5 and 10 lags

OR_cumm_inpat <- OR_cumm(pred.lin, "inpatient")

knitr::kable(OR_cumm_inpat)

 write_csv(OR_cumm_inpat, paste0(output.folder, "inpatients_cumm_lags_OR.csv"))

```
