---
title: "Set years analysis"
author: "Yanelli Nunez"
date: "03/15/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## i.  Loading needed packages

* Objective * 

Construct models for averaged windows of exposure (one model for each window of exposure). The average excluded the 12 months prior to diagnosis (aka year of diagnosis). Models:

- 1 year prior year of hospitalization    
- 5 years prior year of hospitalization
- 10 years prior year of hospitalization

Note
- I ran each of the three models twice, with parish_SES and without parish_SES 
- When I ran the models with parish SES, I lose observations because for about 800 patients/cases I do not have info on parish SES

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages_yanelli.R'))

# load functions
 source(paste0(code.folder, 'models/functions_yanelli.R'))

# load data
 source(paste0(code.folder, 'models/y_00_load_datasets.R'))

```

## ii. Load data and view summary

```{r}

summary(data_1yr)

summary(data_5yrs)

summary(data_10yrs)

```


## 1. Models adjusted for parish-ses


```{r}

## 1-year window of exposure

clogit.1yr.pSES <- survival::clogit(ck ~ PM25_1yr + 
                           family_SES + 
                           place_birth +
                           civst +
                           residence +
                             parish_ses +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_1yr) 


summary(clogit.1yr.pSES)
# extract betas and estimate rate ratio

## beta_trans_clog is a function I created to extract the effect estimates and organize them into a table. You can find this function in the functions_yanelli.R file within the models folder

clog.1yr.est.pSES <- beta_trans_clog(clogit.1yr.pSES, 'clog_1yr')


## 5-year exposure window 

clogit.5yr.pSES <- clogit(ck ~ PM25_5yr + 
                           family_SES + 
                           place_birth +
                           civst +
                       residence +
                       parish_ses +
               strata(nr), # which controls go with which case
              # weights =  
               method = "efron", # the method tells the model how to deal with ties
               data = data_5yrs) 
summary(clogit.5yr.pSES)

# extract betas and estimate rate ratio
clog.5yr.est.pSES <- beta_trans_clog(clogit.5yr.pSES, 'clog_5yr')


## 10-years exposure window 

clogit.10yrs.pSES <- clogit(ck ~ PM25_10yr + 
                           family_SES + 
                           place_birth +
                           civst +
                         residence +
                         parish_ses +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_10yrs) 

summary(clogit.10yrs.pSES)


# extract betas and estimate rate ratio
clog.10yr.est.pSES <- beta_trans_clog(clogit.10yrs.pSES, 'clog_10yr')

```

## 2. Models without parish-ses adjustment

```{r}

## 1-year window of exposure

clogit.1yr <- survival::clogit(ck ~ PM25_1yr + 
                           family_SES + 
                           place_birth +
                           civst +
                           residence +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_1yr) 


summary(clogit.1yr)
# extract betas and estimate rate ratio

## beta_trans_clog is a function I created to extract the effect estimates and organize them into a table. You can find this function in the functions_yanelli.R file within the models folder

clog.1yr.est <- beta_trans_clog(clogit.1yr, 'clog_1yr')


## 5-year exposure window 

clogit.5yr <- clogit(ck ~ PM25_5yr + 
                           family_SES + 
                           place_birth +
                           civst +
                       residence +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_5yrs) 
summary(clogit.5yr)

# extract betas and estimate rate ratio
clog.5yr.est <- beta_trans_clog(clogit.5yr, 'clog_5yr')


## 10-years exposure window 

clogit.10yrs <- clogit(ck ~ PM25_10yr + 
                           family_SES + 
                           place_birth +
                           civst +
                         residence +
               strata(nr), # which controls go with which case
              # weights =  # number of events in each day ?
               method = "efron", # the method tells the model how to deal with ties
               data = data_10yrs) 

summary(clogit.10yrs)


# extract betas and estimate rate ratio
clog.10yr.est <- beta_trans_clog(clogit.10yrs, 'clog_10yr')

```


## 3. Combine estimates from all models 
I'M HERE
```{r}

modl_est <- rbind(clog.1yr.est, clog.5yr.est, clog.10yr.est) %>%
  dplyr::select(model, rr, rr.lci, rr.uci, everything())

knitr::kable(modl_est)

#write_csv(modl_est, paste0(output.folder, "y_04_fix_windows_OR.csv"))
#write_csv(modl_est, paste0(output.folder, "y_04_parish_SES_fix_windows_OR.csv"))


```

