---
title: "Data Statistics Summaries"
author: "Yanelli Nunez"
date: "8/16/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objectives:

Create Table 1 for paper

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))

# load functions
 source(paste0(code.folder, 'models/functions_yanelli.R'))

# load data
 source(paste0(code.folder, 'models/m_00_load_datasets.R'))

```

## i. Number of cases/controls in each data set

Notes: 

- The datasets (1, 5, 10yrs, and lag 10yrs) are largely overlapping so we chose the 5yrs dataset to create the summary statistics of the control/cases (Table 1)
- O refers to controls and 1 to cases

```{r}

ck_1yr <- data_1yr %>%
  group_by(ck) %>%
  summarize(n()) %>%
  mutate(exp_wind = "1 yrs")

ck_5yrs <- data_5yrs %>% 
  group_by(ck) %>%
  summarize(n()) %>%
  mutate(exp_wind = "5 yrs")

ck_10yrs <- data_10yrs %>% 
  group_by(ck) %>%
  summarize(n()) %>%
  mutate(exp_wind = "10 yrs")

ck_lag <- data_lags %>%
  dplyr::select(nr, ck, lobnr) %>%
  unique() %>%
  group_by(ck) %>%
  summarize(n()) %>%
  mutate(exp_wind = "1-yr lags")

ck_full <- rbind(ck_1yr, ck_5yrs, ck_10yrs, ck_lag) %>%
  mutate(ck = as.factor(if_else(ck == "0", "control", "case")))

knitr::kable(ck_full)

```

## 1. Table 1

Since there is a significant amount of overlapping between the datasets, I would use the 5-year average to create table 1

Used Package suggested by Jenni to create tables (it's so helpful!) 
http://www.danieldsjoberg.com/gtsummary/articles/tbl_summary.html

```{r}

table1_prep <- data_5yrs %>%
  mutate(age_diag = trunc((bdate %--% diagdat) / years(1))) %>%
  dplyr::select(ck, sex, age_diag, civst, place_birth, residence, family_SES, PM25_5yr) %>% 
  mutate(ck = as.factor(if_else(ck == "0", "control", "case")),
        sex = as.factor(if_else(sex == "2", "Female", "Male")),
 family_SES = recode_factor(family_SES, 
         `1`='Group 1 (Highest)',
         `2`='Group 2',
         `3`='Group 3',
         `4`='Group 4',
         `5`='Group 5 (Lowest)',
         `9`='Group 9 (Unknown)'),
  civst = recode_factor(civst, 
         `4`='Married',
         `3`='Divorced',
         `2`='Widower',
         `1`='Never married'),
  residence = recode_factor(residence, 
        `Capital Region of Denmark`='Capital Region of Denmark',
        `Cities >50K population`='Cities >50K population',
        `Rest of Denmark`='Rest of Denmark',
        `Unknown`='Unknown'),
  place_birth=recode_factor(place_birth, 
       `copenhagen`='Greater Copenhagen',
       `other_cities`='Big cities of Denmark',
       `other_dk`='Rest of Denmark',
       `greenland`='Greenland',
       `foreign`='Foreign',
       `unknown`='Unknown'))  
 
  
table_1 <- table1_prep %>%
  gtsummary::tbl_summary(#missing = "no",
              label = list(age_diag ~ "Average age (years)",
                           sex ~ "Sex",
                           family_SES ~ "Family SES",
                           place_birth ~ "Place of birth",
                           civst ~ "Civil status",
                           residence ~ "Place of residence",
                           PM25_5yr ~ "PM2.5"),
              by = ck,
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)")) %>%
              add_overall() %>%
  bold_labels()

table_1


```

### i. Save Table 

```{r}
table_1 %>%
  gtsummary::as_flex_table() %>%
  flextable::save_as_docx(path = paste0(tables.folder,'table_1_updated.docx'))
```

