---
title: "Figures"
author: "Yanelli Nunez"
date: "4/21/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

This code you can run because it doesn't use patient data. Here I used the effect estimates I save from the models to make a figure for the effect estimates of average exposures and the distributed lags

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))


```

## 1. Load data 

```{r}


# lag effect estimates from distributed lag models

distrib_lags <- read_csv(paste0(output.folder, "distributed_lags_OR.csv")) %>%
  mutate(lag = factor(lag, levels = c("lag1", "lag2", "lag3", "lag4", "lag5", "lag6", "lag7", "lag8", "lag9", "lag10")),
         lag = recode_factor(lag,       # made names tidyer
                          `lag1` = 'Lag 1',
                          `lag2` = 'Lag 2',
                          `lag3` = 'Lag 3',
                          `lag4` = 'Lag 4',
                          `lag5` = 'Lag 5',
                          `lag6` = 'Lag 6',
                          `lag7` = 'Lag 7',
                          `lag8` = 'Lag 8',
                          `lag9` = 'Lag 9',
                          `lag10` = 'Lag 10'))


# lag Cumulative estimates 

cum_lags <- read_csv(paste0(output.folder, "cumm_lags_OR.csv")) %>%
  dplyr::filter(lag %in% c("lag1", "lag5", "lag10")) %>%
  mutate(exp_wind = as.factor(if_else(lag == "lag1", "1-year",
                               if_else(lag == "lag5", "5-years",
                                       if_else(lag == "lag10", "10-years", "x"))))) %>%
  mutate(exp_wind = factor(exp_wind, levels = c("1-year", "5-years", "10-years"))) %>% 
  dplyr::select(-lag) %>%
  mutate(model = "cumm_exp", parish.SES = if_else(lag_constr == "3-df" | lag_constr == "4-df","Parish SES nonadjusted",                                                   if_else(lag_constr == "3-df with/parish ses", "Parish SES adjusted",
                                          if_else(lag_constr == "3-df w/out parish ses/indiv", "W/out Parish SES/individ", lag_constr)))) %>%
  dplyr::filter(lag_constr != "4-df") %>%
  dplyr::select(-lag_constr)

# fix windows 

fix_windw <- read_csv(paste0(output.folder, "fix_windows_OR.csv")) %>%
  mutate(model.clean = as.factor(if_else(model == "clog_1yr", "1-year",
                               if_else(model == "clog_5yr", "5-years",
                                       if_else(model == "clog_10yr", "10-years", "x"))))) %>%
  mutate(model.clean = factor(model.clean, levels = c("1-year", "5-years", "10-years")))  %>% 
  dplyr::select(-beta, -beta.se, -model) %>%
  dplyr::rename(OR_cumm = rr,
                lower_CI = rr.lci,
                upper_CI = rr.uci,
               # model = parish.SES,
                exp_wind = model.clean) %>%
   mutate(model = "avgr_exp")


# Combined cumulative lags with effect estimates from averaged exposure
cum_exp_full <- rbind(cum_lags, fix_windw) %>%
  mutate(parish.SES = as.factor(parish.SES),
         model = as.factor(model))
  

```


## 2. Forest plots

### 2a. Sensitivity analysis for parish-level SES 

- Average windows of exposure and cumulative lag effects 
- This is figure 1 in the manuscript and the results are from the models adjusted for parish SES
- For the cumulative effect estimates, the lag constrain is 3df


```{r}

pd <- position_dodge(width = 0.5)

averg_cum_expo_SES <- cum_exp_full %>%
dplyr::filter(parish.SES %in% c("W/out Parish SES/individ", "Parish SES adjusted")) %>%
  ggplot(aes(x = exp_wind, y = OR_cumm, ymin = lower_CI, ymax = upper_CI)) +
   geom_hline(yintercept = 1, linetype = 1) +
    geom_pointrange(aes(color = model, linetype = parish.SES), position = pd, size = 0.5) +
    ylab("Odds ratio per"~mu*g/m^3~increase) +
    xlab("Exposure time window (pre-diagnosis)") +
   #geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = pd, width = 0.20) +
 #coord_flip()  +
 # scale_y_continuous(limits = c(0.95, 1.10), breaks = seq(.95, 1.10, by = 0.05)) +
  theme_minimal() +
    theme(
        axis.ticks.y=element_blank(),
        axis.text.x = element_text(size = 17),
        axis.title.x = element_text(vjust = - 0.3),
        axis.text.y = element_text(size = 17),
        axis.title = element_text(size = 18),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 17),
        strip.text = element_text(size = 18),
        legend.key.size = unit(2.5, "lines")) +
  scale_color_manual(labels = c("Averg. Exposure", "Cum. Exposure"), values = c("black", "#1f78b4")) +
  scale_linetype_manual(labels = c("Parish-SES adjusted", "Parish-SES not adjusted"), values = c("solid", "dashed"),
                         guide = guide_legend(override.aes = list(shape = NA)))


#  guides(linetype = guide_legend(override.aes = list(linetype = c("solid", "dashed"))))
  #  guides(linetype = guide_legend(override.aes = list(shape = c(0, 0) ) ) )
 # guides(colour = guide_legend(override.aes = list(size = 10))


 
averg_cum_expo

```


### 2b. Sensitivity analysis for including/excluding individuals missing parish SES

This figure is for the sensitivity analysis and will be included in the supplements. It shows the lag cumulative effects and the effect estimates of average exposures from the models that were *NOT* adjusted for Parish SES

```{r}

pd <- position_dodge(width = 0.5)


averg_cum_expo_missing <- cum_exp_full %>%
dplyr::filter(parish.SES %in% c("Parish SES nonadjusted", "W/out Parish SES/individ")) %>%
  ggplot(aes(x = exp_wind, y = OR_cumm, ymin = lower_CI, ymax = upper_CI)) +
   geom_hline(yintercept = 1, linetype = 1) +
    geom_pointrange(aes(color = model, linetype = parish.SES), position = pd, size = 0.5) +
    ylab("Odds ratio per"~mu*g/m^3~increase) +
    xlab("Exposure time window (pre-diagnosis)") +
   #geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = pd, width = 0.20) +
 #coord_flip()  +
 # scale_y_continuous(limits = c(0.95, 1.10), breaks = seq(.95, 1.10, by = 0.05)) +
  theme_minimal() +
    theme(
        axis.ticks.y=element_blank(),
        axis.text.x = element_text(size = 17),
        axis.title.x = element_text(vjust = - 0.3),
        axis.text.y = element_text(size = 17),
        axis.title = element_text(size = 18),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 17),
        strip.text = element_text(size = 18),
        legend.key.size = unit(2.5, "lines")) +
  scale_color_manual(labels = c("Averg. Exposure", "Cum. Exposure"), values = c("gray24", "darkgreen")) +
  scale_linetype_manual(labels = c("Full dataset", "Data subset"), values = c("solid", "dashed"),
                         guide = guide_legend(override.aes = list(shape = NA)))
 
  

averg_cum_expo_unadjust
```

## 3. Distributed lag figure

### 3b. 3df in the lag constrain 

This figure compares the distributed lag results from including parish SES vs excluding parish SES

```{r}


distr_lag_pSES <- distrib_lags %>%
  dplyr::filter(lag_constr %in% c("3-df w/parish ses", "3-df w/out parish ses/indiv")) %>%
   mutate(lag_constr = factor(lag_constr, levels = c("3-df w/parish ses", "3-df w/out parish ses/indiv"))) %>%
  ggplot(aes(x = lag, y = OR, ymin = lower_CI, ymax = upper_CI)) +
   geom_hline(yintercept = 1, linetype = 1) +
   geom_pointrange(aes(linetype = lag_constr), position = pd, size = 0.5) +
    ylab("Odds ratio per"~mu*g/m^3~increase) +
    xlab("Exposure time window (pre-diagnosis)") +
    #geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = pd, width = 0.20) +
  scale_y_continuous(limits = c(0.95, 1.10), breaks = seq(.95, 1.10, by = 0.05)) +
  theme_minimal() +
    theme(
        axis.ticks.y=element_blank(),
        axis.text.x = element_text(size = 17, angle = 75),
        axis.title.x = element_text(vjust = - 0.2),
        axis.text.y = element_text(size = 17),
        axis.title = element_text(size = 18),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 17),
        strip.text = element_text(size = 18),
        legend.key.size = unit(2.5, "lines"))  +
   scale_linetype_manual(labels = c("Parish-SES adjusted", "Parish-SES not adjusted"), values = c("solid", "dashed"),
                         guide = guide_legend(override.aes = list(shape = NA)))


distr_lag


```

### 3c. 3df in lag constrain models including/excluding participants with missing parish-level SES



```{r}

distr_lag_missing <- distrib_lags %>%
  dplyr::filter(lag_constr %in% c("3-df", "3-df w/out parish ses/indiv")) %>%
   mutate(lag_constr = factor(lag_constr, levels = c("3-df", "3-df w/out parish ses/indiv"))) %>%
  ggplot(aes(x = lag, y = OR, ymin = lower_CI, ymax = upper_CI)) +
   geom_hline(yintercept = 1, linetype = 1) +
   geom_pointrange(aes(linetype = lag_constr), position = pd, size = 0.5) +
    ylab("Odds ratio per"~mu*g/m^3~increase) +
    xlab("Exposure time window (pre-diagnosis)") +
    #geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = pd, width = 0.20) +
  scale_y_continuous(limits = c(0.95, 1.10), breaks = seq(.95, 1.10, by = 0.05)) +
  theme_minimal() +
    theme(
        axis.ticks.y=element_blank(),
        axis.text.x = element_text(size = 17, angle = 75),
        axis.title.x = element_text(vjust = - 0.2),
        axis.text.y = element_text(size = 17),
        axis.title = element_text(size = 18),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 17),
        strip.text = element_text(size = 18),
        legend.key.size = unit(2.5, "lines"))  +
   scale_linetype_manual(labels = c("Full dataset", "Data subset"), values = c("solid", "dashed"),
                         guide = guide_legend(override.aes = list(shape = NA)))

```

## 4. Sex stratified analysis

### 4b. Organize data


```{r}

# lag effect estimates from distributed lag models

sex_lags <- read_csv(paste0(output.folder, "sex_distributed_lags_OR.csv")) %>%
  mutate(lag = factor(lag, levels = c("lag1", "lag2", "lag3", "lag4", "lag5", "lag6", "lag7", "lag8", "lag9", "lag10")),
         lag = recode_factor(lag,       # made names tidyer
                          `lag1` = 'Lag 1',
                          `lag2` = 'Lag 2',
                          `lag3` = 'Lag 3',
                          `lag4` = 'Lag 4',
                          `lag5` = 'Lag 5',
                          `lag6` = 'Lag 6',
                          `lag7` = 'Lag 7',
                          `lag8` = 'Lag 8',
                          `lag9` = 'Lag 9',
                          `lag10` = 'Lag 10'))


# lag Cumulative estimates 

sex_cum_lags <- read_csv(paste0(output.folder, "sex_cumm_lags_OR.csv")) %>%
  dplyr::filter(lag %in% c("lag1", "lag5", "lag10")) %>%
  mutate(exp_wind = as.factor(if_else(lag == "lag1", "1-year",
                               if_else(lag == "lag5", "5-years",
                                       if_else(lag == "lag10", "10-years", "x"))))) %>%
  mutate(exp_wind = factor(exp_wind, levels = c("1-year", "5-years", "10-years"))) %>% 
  dplyr::select(-lag)

```

### 4c. Sex stratified figure 

```{r}

# cummulative lag effects
pd <- position_dodge(width = 0.5)


sex_fig_cum_lags <- sex_cum_lags %>%
  ggplot(aes(x = exp_wind, y = OR_cumm, ymin = lower_CI, ymax = upper_CI)) +
   geom_hline(yintercept = 1, linetype = 1) +
    geom_pointrange(aes(linetype = model), position = pd, size = 0.5) +
    ylab("") +
    xlab("") +
  ggtitle("(B) Cumulative effect estimates") +
   #geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = pd, width = 0.20) +
 #coord_flip()  +
 # scale_y_continuous(limits = c(0.95, 1.10), breaks = seq(.95, 1.10, by = 0.05)) +
  theme_minimal() +
    theme(
        axis.ticks.y=element_blank(),
        axis.text.x = element_text(size = 17),
        axis.title.x = element_text(vjust = - 0.3),
        axis.text.y = element_text(size = 17),
        axis.title = element_text(size = 18),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 17),
        strip.text = element_text(size = 18),
        legend.key.size = unit(2.5, "lines"),
         plot.title = element_text(size=18)) +
  scale_linetype_manual(labels = c("Female", "Male"), values = c("solid", "dashed") )
 
sex_fig_cum_lags

# lag effects

sex_fig_lags <- sex_lags %>%
  ggplot(aes(x = lag, y = OR, ymin = lower_CI, ymax = upper_CI)) +
   geom_hline(yintercept = 1, linetype = 1) +
   geom_pointrange(aes(linetype = model), position = pd, size = 0.5) +
    ylab("") +
    xlab("") +
  ggtitle("(A) Effect estimates per lag") +
    #geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), position = pd, width = 0.20) +
  scale_y_continuous(limits = c(0.93, 1.13), breaks = seq(.95, 1.10, by = 0.05)) +
  theme_minimal() +
    theme(
        axis.ticks.y=element_blank(),
        axis.text.x = element_text(size = 17, angle = 75),
        axis.title.x = element_text(vjust = - 0.2),
        axis.text.y = element_text(size = 17),
        axis.title = element_text(size = 18),
        legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 17),
        strip.text = element_text(size = 18),
        legend.key.size = unit(2.5, "lines"),
        plot.title = element_text(size=18))  +
   scale_linetype_manual(labels = c("Female", "Male"), values = c("solid", "dashed"),
                         guide = guide_legend(override.aes = list(shape = NA)))


# combine figures 

yleft <- textGrob("Odds ratio per"~mu*g/m^3~PM[2.5]~increase, rot = 90, gp = gpar(fontsize = 20))

grid.arrange(sex_fig_lags,
             sex_fig_cum_lags,
             nrow = 2, ncol = 1, left = yleft)


```

