---
title: "Mergin air pollution & patient data"
author: "Yanelli"
date: "8/04/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective: 

Combine the pollution and patient data into a single dataset


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))

```

## 1. Loading data

### 1a. Load pollution data

- Pollution data are in monthly averages
- There are two different pollution measures in the dataset but the one we are interested in is the one representing street level concentrations. The street level pollution variables have a _str suffix 

```{r}

# Load pollution data
load(paste0(raw_data_files, "ALS_Final_pol_1989-2013_1989_5_2k.RData"))

# rename file
pollutant_time_series <- `ALS_Final_pol_1989-2013_1989_5_2k`

# check how many missing observations each variable has
missing <- as.data.frame(colSums(is.na(pollutant_time_series)))

summary(pollutant_time_series)

# select "str" pollutants
pollution_data <- pollutant_time_series %>%
  mutate(date_exp = make_date(year, month)) %>% # merge month and year of exposure into single variable
  dplyr::select(lobnr, nr, date_exp, PM25_str, CO_str, EC_str, NO2_str, NOX_str, O3_str, PM10_str) # select pollutants of interest

```

### 1b. Load patient data 

- This is the data that with the summarize categorical variables
- I created variables that represent the time limits for each window of exposure. This facilitates calculating the exposure average for each case/control 

```{r}
# load patient data
# create the time limits for the windows of exposure
patient_data <- readRDS(paste0(raw_data_files, "patient_data_uniform_complete.rds"), NULL) %>%
  mutate(year_gap = diagdat - years(1), # one year gap pre diagnosis
         one_year = year_gap - years(1), # date limit for 1 year average
         five_years = year_gap - years(5), # date limit for 5 year average
         ten_years = year_gap - years(10)) %>%  #date limit for 10 year average
  dplyr::select(nr, lobnr, ck, diagtype, diagdat, year_gap, one_year, five_years, ten_years, everything()) %>%   
  dplyr::select(-daars3, -daars4, -daars5) %>% # I won't need this variables
  dplyr::rename(family_SES = famses) %>% # this is just a personal preference in the naming style
  mutate(birth_year = year(bdate)) # create a variable for year of birth

summary(patient_data)

# Check how many cases I have total
cases <- patient_data %>%
  dplyr::filter(ck == 1) %>%
 dplyr::group_by(nr) %>%
  summarise(count = n())


# Check how many controls I have total
contrl <- patient_data %>%
 dplyr:: filter(ck == 0) %>%
  dplyr::group_by(nr, lobnr) %>%
  summarise(count = n())


```



## 2. Bind hospitalization and exposure date

```{r}

full_dataset <- left_join(patient_data, pollution_data, by = c("nr", "lobnr")) %>%
  ungroup()

summary(full_dataset)

# check number of cases again
cases_full <- full_dataset %>%
  dplyr::filter(ck == 1) %>%
 dplyr::group_by(nr) %>%
  summarise(count = n())


# check number of controls again
contrl_full <- full_dataset %>%
 dplyr:: filter(ck == 0) %>%
  dplyr::group_by(nr, lobnr) %>%
  summarise(count = n())

```

## 3. Save dataset 

```{r}
# save data
saveRDS(full_dataset, paste0(yanelli_datasets, "joined_datasets.rds")) #save file of hospitalizations and pollution joined dataset
```

