---
title: "Windows_exposure"
author: "Yanelli Nunez"
date: "08/05/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective: 
- To select the months of exposure that follow within the exposure windows periods (1, 5, and 10 years pre-hospitalization). This is done in the individual level and years are estimated from diagnosis date.

- We also set some criteria for the cases/controls that should be included in the analysis. More details about this are presented below. 


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))

```

## 1. load merged patient/pollution dataset 
```{r}

full_dataset <- readRDS(paste0(yanelli_datasets, "joined_datasets.rds")) # this is a large dataset

summary(full_dataset)

```


## 2. Nest data by cases 

This allow us to more easily create windows of exposure for each case

```{r}

# a data set made up of sub-dataset. Each data subset corresponds to a patient and his/her controls
nested_data <- full_dataset %>% 
  mutate(nr = as.factor(nr),
         lobnr = as.factor(lobnr),
         ck = as.factor(ck)) %>%
  group_by(nr) %>%
  nest()

summary(nested_data)

### take a look in a single case and corresponding controls 
test <- as.data.frame(nested_data$data[[1]]) %>%
 dplyr::filter(date_exp <= year_gap & date_exp >= one_year)

summary(test)
  

```


## 3. Select exposure within window of exposure period

- For each case, select years within exposure window of 1, 5, and 10 years pre-hospitalization with a 1-year gap between start of exposure and first hospitalization

```{r}

# filter data for the 1, 5 and 10 year periods

filtered_data <- nested_data %>%
  mutate(one_year = purrr::map(data, ~ dplyr::filter(.x, date_exp <= year_gap & date_exp > one_year)),
         five_yr = purrr::map(data, ~ dplyr::filter(.x, date_exp <= year_gap & date_exp > five_years)),
         ten_yr = purrr::map(data, ~ dplyr::filter(.x, date_exp <= year_gap & date_exp > ten_years)))
 
# explore a random dataset within the list of datasets to make sure it looks right
explore <- filtered_data$one_year[[1]]
summary(explore)


```


## 4. Explore missing data 

### 4a. One-year exposure window

Notes: 

1) In the original dataset there is a total of 4,011 cases. If each case has 5 controls then there should be a total of 20,055 controls, that is, 24,066 net total. 

2) The one-year window of exposure has a total of 23,732 case/controls with a **total of 3,952 cases** specifically and ***19,780 controls***. This means **59 cases** from the original data have zero observations during this year. That is, the data is missing ***1.5%*** of cases from the original dataset

3) Overall 93 out of 23,732 cases/controls **(0.39%)** are missing one or more of the 12 months for the one year exposure window


```{r}

one_year <- filtered_data %>%
  dplyr::select(-data, -five_yr, -ten_yr) %>%
  unnest(., cols = one_year)

# check how many cases and controls total we have in this data set
case_num <- one_year %>%
  dplyr::filter(ck == 1) %>% # selecting only cases
  count(nr) # count number of unique "nr" (cases)

control_num <- one_year %>%
 dplyr::filter(ck == 0) %>% # selecting controls
  group_by(nr) %>% 
   count(lobnr)

# check how many cases/controls are missing months
exp_missing <- one_year %>%
 group_by(nr, lobnr) %>%
  summarise(counts = n()) # lowest number of months measure for a given case or control is 1

exp_missing %>%
  tally(counts<12) %>%
  summarize(n=sum(n))

# heat map to visualize missingness
exp_missing %>% ggplot(aes(as.numeric(nr), lobnr, fill= counts)) + 
  geom_tile() +
  theme_classic() +
  scale_fill_gradient('counts', limits=c(1, 12), breaks = c(1, 3, 6, 9, 12),  low = "red", high = "darkblue") 

```

## 4b. 5-year exposure window

Notes: 

1) In the original dataset there is a total of 4011 cases. If each case has 5 controls then they should be a total of 20,055 controls, that is, 24,066 net total

2) The 5-year window of exposure has a total of 23,801 cases/controls. Specifically, it has **3,965 cases** and **19,836 controls**. This means **46 cases** from the original data have zero observations during this 5-year time period. That is, the data is missing ***1.1%*** of cases from the original dataset

3) 434 out of 23,801 cases/controls **(~1%)** are missing one or more of the 60 months of the five year exposure window


```{r}

# select the 5-years time window dataset
five_years <- filtered_data %>%
  dplyr::select(-data, -one_year, -ten_yr) %>%
  unnest(., cols = five_yr)

# check how many cases total we have in this dataset
case_num_5 <- five_years %>%
 dplyr::filter(ck == 1) %>% # selecting only cases
  count(nr) # count number of unique "nr" (cases)

# check how many controls total we have in this dataset
control_num_5 <- five_years %>%
  dplyr::filter(ck == 0) %>% # selecting controls
  group_by(nr) %>% 
   count(lobnr)

# check how many cases/controls are missing months
exp_missing_5yrs <- five_years %>%
  group_by(nr, lobnr) %>%
  summarise(counts = n()) # minimum number of available months is 1

exp_missing_5yrs %>%
  tally(counts<60) %>%
  summarize(n=sum(n))

# heat map to visualize missingness
exp_missing_5yrs %>% ggplot(aes(as.numeric(nr), lobnr, fill= counts)) + 
  geom_tile() +
  theme_classic() +
  scale_fill_gradient('counts', limits=c(1, 60), breaks = c(1, 12, 24, 36, 48, 60),  low = "red", high = "darkblue") 


```


## 4c. 10-year exposure window

Notes: 

1) In the original dataset there is a total of 4,011 cases. If case has 5 controls then there should be a total of 20,055 controls, that is, 24,066 net total

2) The 10-year window of exposure has a total of 23,848 cases/controls. Specifically, it has **3,974 cases** and ***19,874 controls***. This means **37 cases** from the original data have zero observations during this 10-year time period. That is, the data is missing ***0.9%*** of cases from the original dataset

3) 1284 out of 23,848 cases/controls **(~5%)** are missing one or more of the 120 months of the 10-year exposure period


```{r}

# select the ten-years time window dataset
ten_years <- filtered_data %>%
  dplyr::select(-data, -one_year, -five_yr) %>%
  unnest(., cols = ten_yr)

# check how many cases total we have in this dataset
case_num_10 <- ten_years %>%
 dplyr::filter(ck == 1) %>% # selecting only cases
  count(nr) # count number of unique "nr" (cases)

# check how many controls total we have in this dataset
control_num_10 <- ten_years %>%
  dplyr::filter(ck == 0) %>% # selecting controls
  group_by(nr) %>% 
   count(lobnr)

# check how many cases/controls are missing months
exp_missing_10yrs <- ten_years %>%
  group_by(nr, lobnr) %>%
  summarise(counts = n()) # minimum number of available months is 1

exp_missing_10yrs %>%
  tally(counts<120) %>%
  summarize(n=sum(n))

# heat map to visualize missingness
exp_missing_10yrs %>% ggplot(aes(as.numeric(nr), lobnr, fill= counts)) + geom_tile() +
  #scale_fill_continuous(limits=c(0, 12), breaks=seq(0,12, by=1)) +
  #scale_fill_viridis(limits = c(0, 60), discrete=FALSE) +
  theme_classic() +
  scale_fill_gradient('counts', limits=c(1, 120), breaks = c(1, 24, 48, 72, 96, 120),  low = "red", high = "darkblue") 

```

## 5. Data selection

Criteria:

For the 1-year exposure window, limit cases and controls to those we have at least 75% of the pollutant measurements. That is, at least **9 out of the 12 months** need to have a pollutant measurement. Furthermore, within those 9 measurements we need to have at least one measure for each of the year seasons (Denmark has 4 seasons) 

For the 5-years and 10-years exposure window, limit cases and controls to those we have at least 50% of the PM2.5 measurements. That is, for the 5-year time window at least **30 out of the 60 months** should have a measurement. For the 10-year window, each case/control should have at least 60 months with pollutant measurements out of the 120 months. 

For all time windows, each case needs to have at least one control that fulfills the requirements described in the above text. 



### 5a. 1-year exposure windows

Notes: 

- **3,942 cases and 19,406 controls total** after selecting only the data that meets the requirements described above
Now each case should have at least one control and each entry should have pollutant measurements for at least 9 out of the 12 months and there should be at least one measurement in each of the four weather seasons


```{r}

filter_1yr <- one_year %>%
  group_by(nr, lobnr) %>%
  dplyr::mutate(months_meas = n()) %>%
  ungroup() %>%
  dplyr::filter(months_meas >= 9) %>% # select those with at least 9 months 
  mutate(month_exp = month(date_exp),
         season = if_else(month_exp >=3 & month_exp <= 5, "spring", 
                  if_else(month_exp >=6 & month_exp <=8, "summer",
                  if_else(month_exp >=9 & month_exp <= 11, "fall",
                  if_else(month_exp == 12 | month_exp <= 2, "winter", "x"))))) %>%  # season ID
  group_by(nr, lobnr) %>%
  mutate(season_num = length(unique(season))) %>% #XX entries are missing measurements in a season
ungroup() %>%
  dplyr::filter(season_num >= 4) %>% # filter for entries that have at least 1 measurement in each season
group_by(nr) %>%
  dplyr::mutate(case_control = length(unique(ck))) %>% # ID to check that each case has at least one control and that each control has a case 
  dplyr::filter(case_control >= 2) %>% # filter cases that have  at least  one  control  and controls that have a  case
  ungroup() %>%
  dplyr::select(-month_exp, -months_meas, -season, -season_num, -case_control) # remove IDs
         
summary(filter_1yr)


# final number of cases
length(unique(filter_1yr$nr))

#final number of controls
filter_1yr %>%
  dplyr::filter(ck == 0) %>% #select controls only
  group_by(nr) %>%
 count(lobnr) %>%
  nrow()
    
# save data 
write_csv(filter_1yr, paste0(yanelli_datasets, "one_yr_filtered_data.csv"))

```


### 5b. 5-years exposure windows

Notes: 

- **3,942 cases and 19,398 controls** total after selecting only the data that meets the requirements 

```{r}

filter_5yrs <- five_years %>%
  group_by(nr, lobnr) %>%
  dplyr::mutate(months_meas = n()) %>% #ID to check mounts measure
  ungroup() %>%
  dplyr::filter(months_meas >= 30) %>% # select those with at least 30 months 
group_by(nr) %>%
  dplyr::mutate(case_control = length(unique(ck))) %>% # ID to check that each case has at least one control and that each control has a case 
  dplyr::filter(case_control >= 2) %>%
  ungroup() %>%
  dplyr::select(-case_control, -months_meas) # remove IDs
         
summary(filter_5yrs)

# final number of cases
length(unique(filter_5yrs$nr))

#final number of controls
filter_5yrs %>%
  dplyr::filter(ck == 0) %>% #select controls only
  group_by(nr) %>%
 count(lobnr) %>%
  nrow()
    

# save data 
write_csv(filter_5yrs, paste0(yanelli_datasets, "five_yrs_filtered_data.csv"))

```

### 5c. 10-years exposure window 

Notes: 
- **3,938 cases and 19,362 controls** total after selecting only the data that meets the requirements

```{r}

filter_10yrs <- ten_years %>%
  group_by(nr, lobnr) %>%
  dplyr::mutate(months_meas = n()) %>% #ID to check months measure
  ungroup() %>%
  dplyr::filter(months_meas >= 60) %>% # select those with at least 60 months 
group_by(nr) %>%
  dplyr::mutate(case_control = length(unique(ck))) %>% # ID to check that each case has at least one control and that each control has a case 
  dplyr::filter(case_control >= 2) %>%
  ungroup() %>%
  dplyr::select(-case_control, -months_meas) # remove ID
         
summary(filter_10yrs)

# final number of cases
length(unique(filter_10yrs$nr))

#final number of controls
filter_10yrs %>%
 dplyr::filter(ck == 0) %>% #select controls only
  group_by(nr) %>%
 count(lobnr) %>%
  nrow()
    

# save data 
write_csv(filter_10yrs, paste0(yanelli_datasets, "ten_yrs_filtered_data.csv"))

```

