---
title: "Distributed lag"
author: "Yanelli Nunez"
date: "2/5/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:

To create yearly exposure average back to 10 years from date of diagnosis (individually for each case and control) 


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))

```

## 1. Load dataset 

Notes: 

Total cases = 4011
Total controls = 20055

1600 observations missing air pollution data (0.02% observations)

```{r}

full_dataset <- readRDS(paste0(yanelli_datasets, "joined_datasets.rds")) %>%
  dplyr::select(-year_gap, -one_year, -five_years, -ten_years) %>% # remove variables we don't need
  mutate(ten_years = diagdat - lubridate::years(10),
         first_year = diagdat - lubridate::years(1),
         second_year = first_year - lubridate::years(1)) %>% # date limit for 10 years back from diagnosis date
  dplyr::select(nr, lobnr, ck, diagtype, diagdat, ten_years, everything())

summary(full_dataset)

# number of cases and controls 
# cases 
total_cases <- full_dataset %>%
  dplyr::filter(ck == 1) %>%
 dplyr::group_by(nr) %>%
  summarise(count = n())

count(total_cases)

#controls
total_contrl <- full_dataset %>%
 dplyr::filter(ck == 0) %>%
  dplyr::group_by(nr, lobnr) %>%
  summarise(count = n())

count(total_contrl)


```

## 2. Nest data

- Nest data by cases so we can easily select 10 years of exposure for each case/control
- The nested dataset is made up of sub-dataset. Each data subset corresponds to a patient and his/her controls
```{r}

# nest data
nested_data <- full_dataset %>% 
  mutate(nr = as.factor(nr),
         lobnr = as.factor(lobnr),
         ck = as.factor(ck)) %>%
  group_by(nr) %>%
  tidyr::nest()



### take a look in a single case and corresponding controls 
test <- as.data.frame(nested_data$data[[2]]) %>%
 dplyr::filter(date_exp <= diagdat & date_exp >= ten_years)

summary(test)
  
```


## 3. Select exposure during the 10 year period

For each case/control, filter 10 years of exposure from date of hospitalization  

```{r}

filtered_data <- nested_data %>%
  mutate(ten_yrs = purrr::map(data, ~ dplyr::filter(.x, date_exp <= diagdat & date_exp > ten_years)))
  

# explore a random dataset within the list of datasets to make sure it looks right
explore <- filtered_data$ten_yrs[[2]]
summary(explore)


# unnest dataset for 10 years 
ten_year <- filtered_data %>%
  dplyr::select(-data) %>%
  unnest(., cols = ten_yrs)


```


```{r}

review <- ten_year %>%
  mutate(year_exp = year(date_exp)) %>%
 group_by(nr, lobnr, year_exp) %>%
  mutate(months_meas = n()) %>%
  ungroup() %>%
  group_by(nr, lobnr, year_exp) %>%
 dplyr::select(nr, lobnr, year_exp, months_meas) %>%
  unique() %>%
  ungroup() %>%
 group_by(nr, lobnr, months_meas) %>%
  summarise(numb_years = n()) %>%
  ungroup()  %>%
dplyr::filter(months_meas == 1) %>%
  dplyr::select(nr, lobnr) %>%
  unique()

review

```


## 4. Annual averages 

I estimated annual averages of exposure for the six pollutants we have data for. I'm only working with PM2.5 but Robbie use the other pollutants for his analysis


```{r}

# averaging 
avgr <- ten_year %>%
  mutate(year_exp = year(date_exp)) %>%
  dplyr::select(-PM10_str) %>% #remove because otherwise the following step leads to so many observations that the rest in the pipe fails--we don't need pm10 anyways
  gather(., pollutant, concentration, 32:37) %>%
  group_by(nr, lobnr, pollutant, year_exp) %>%
  mutate(mean_conc = mean(concentration)) %>%
  dplyr::select(-concentration, -date_exp) %>%
  dplyr::distinct() %>%
  spread(., pollutant, mean_conc) %>%
  ungroup()

unique(avgr$pollutant)

summary(avgr)

 write_csv(avgr, paste0(yanelli_datasets, "lagged_10yrs.csv")) # your dataset contains the updated file
avgr <- read_csv(paste0(yanelli_datasets, "lagged_10yrs.csv"))


```

# Visualize data

Notes: 
-We have a total of **3981 cases** and **19,889** controls in the 10-year lagged by year data
-3603 cases have measurements for all years (90.50%)
-17,991 controls have measurements for all years (90.51%)
- 258,476 observations total 



```{r}

# check how many cases total we have in this dataset

avgr_cases <- avgr %>%
 dplyr::filter(ck == 1) %>% # selecting only cases
    dplyr::group_by(nr) %>% 
  dplyr::count(nr) # count number of unique "nr" (cases)

summary(avgr_cases)

# check how many controls total we have in this dataset

avgr_contrl <- avgr %>%
  dplyr:: filter(ck == 0) %>% # selecting controls
  dplyr::group_by(nr) %>% 
  dplyr:: count(lobnr)

summary(avgr_contrl)

# heat map to visualize missingness

avgr_viz <- avgr %>% 
  group_by(nr,lobnr) %>%
  summarise(counts = n())

 
 avgr_viz %>% ggplot(aes(as.numeric(nr), as.factor(lobnr), fill= counts)) + 
  geom_tile() +
  #scale_fill_continuous(limits=c(0, 12), breaks=seq(0,12, by=1)) +
  #scale_fill_viridis(limits = c(0, 60), discrete=FALSE) 
  theme_classic() +
  scale_fill_gradient('counts', limits=c(1, 11), breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),  low = "red", high = "darkblue")


# missingess distribution
avgr_viz %>% ggplot(aes(as.factor(counts))) +
    geom_histogram(stat = "count") 

```

# select only the cases that have measurenments for all 10 years backwards from diagnosis 

Notes:
We have *17991* complete controls (annual pm2.5 measurement wise)
We have *3603* complete cases (annual pm2.5 measurement wise)

All case have at least 2 controls



OLD DATA
We have 8549 complete controls (annual pm2.5 measurement wise)
We have 1714 complete cases (annual pm2.5 measurement wise)

85 cases don't have any controls
384 controls don't have any cases

```{r}

complete_avgr <- avgr %>%
  dplyr:: group_by(nr,lobnr) %>%
 dplyr::mutate(counts = n()) %>%
  dplyr::filter(counts == 11) %>%
  dplyr::select(-counts) 
  

write_csv(complete_avgr, paste0(yanelli_datasets, "lagged_10yrs_complete.csv")) # your dataset folder contains the updated file

# check how many cases and controls we ended up with

num_controls <- complete_avgr %>%
dplyr::filter(ck == 0) %>% # selecting controls
 group_by(nr) %>% 
  count(lobnr)

num_cases <- complete_avgr %>%
dplyr::filter(ck == 1) %>% # selecting only cases
 count(nr) 

#number of controls each case has

cont_case <- complete_avgr %>%
  dplyr::group_by(nr, lobnr) %>%
  dplyr::summarize() %>%
  ungroup() %>%
  dplyr::group_by(nr) %>%
  mutate(numb = n())
  
```

