---
title: "Lag analysis"
author: "Yanelli Nunez"
date: "9/29/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
The SES indicator is based on the proportion (2009) of highest attained educational level in a parish (the smallest administrative unit level in Denmark) of inhabitants aged 30 to 60 years old. We unfortunately only have this information for the year (2009).

SES variable dictionary (numbers in each column represent percentage:
ses0 = unknown
ses1 = Basic
ses23 = II High School
ses4 = III vocational
ses567 = IV short to middle long academic training
ses89 = V university
sesto = total sum of percentage

For using as an SES variable we added SEs for II high school and above

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'file_locations.R'))

# load packages
 source(paste0(packages.folder,'packages.R'))


```

## 1. Load data

- I removed cases/controls with Greenland residence because there is only very few of them 

- I also removed cases/controls with missing civic status (civist) since this is one of the covariants we use in the models 

- After removing the above, I checked that all cases still have at least one control

- I converted all categorical variables to factors

```{r}

# 1 year

data_1yr <- read_csv(paste0(yanelli_datasets, "one_yr_avg_data.csv")) %>%
 mutate(family_SES = as.factor(family_SES),
        sex = as.factor(sex),
        geokode = as.factor(geokode),
        civst = as.factor(civst),
        place_birth = as.factor(place_birth),
        residence = as.factor(residence)) %>%
# dplyr::filter(residence != "Greenland") %>% already remove in a previous .rmd
   droplevels() %>%
 # tidyr::drop_na(civst) %>% # already remove in a previous .rmd
  group_by(nr) %>% 
  mutate(sum=sum(ck)) %>% 
  dplyr::filter(sum==1) %>%
   dplyr::select(-sum) %>%
  ungroup() 

data_1yr %>%
  mutate(ck = as.factor(ck)) %>%
  summary()

class(data_1yr)


# 5 years

data_5yrs <- read_csv(paste0(yanelli_datasets, "five_yrs_avg_data.csv")) %>%
  mutate(family_SES = as.factor(family_SES),
         sex = as.factor(sex),
         geokode = as.factor(geokode),
         civst = as.factor(civst),
         place_birth = as.factor(place_birth),
         residence = as.factor(residence)) %>%
 # dplyr::filter(residence != "Greenland") %>%
  droplevels() %>%
 # tidyr::drop_na(civst) %>%
  group_by(nr) %>% mutate(sum=sum(ck)) %>% 
  dplyr::filter(sum==1) %>% # keeping only the controls that have a case
   dplyr::select(-sum) %>%
  ungroup() 

data_5yrs %>%
  mutate(ck = as.factor(ck)) %>%
  summary()

# 10 years

data_10yrs <- read_csv(paste0(yanelli_datasets, "ten_yrs_avg_data.csv"))%>%
  mutate(family_SES = as.factor(family_SES),
         sex = as.factor(sex),
         geokode = as.factor(geokode),
         civst = as.factor(civst),
         place_birth = as.factor(place_birth),
         residence = as.factor(residence)) %>%
 # dplyr::filter(residence != "Greenland") %>%
   droplevels() %>%
#  tidyr::drop_na(civst) %>%
  group_by(nr) %>% mutate(sum=sum(ck)) %>%
  dplyr::filter(sum==1) %>% #keeping only the controls that have a case
  dplyr::select(-sum) %>%
  ungroup()

data_10yrs %>%
  mutate(ck = as.factor(ck)) %>%
  summary()

# lags
lag_10yrs <- read_csv(paste0(yanelli_datasets, "lagged_10yrs.csv")) %>%
  mutate(family_SES = as.factor(family_SES),
         place_birth = as.factor(place_birth),
         residence = as.factor(residence),
         civst = as.factor(civst)) %>%
  # dplyr::filter(residence != "Greenland") %>%
   droplevels() %>%
 # tidyr::drop_na(civst) %>%
  group_by(nr) %>% mutate(sum=sum(ck)) %>% 
  dplyr::filter(sum==11) %>% # keeping only the controls that have a case
   dplyr::select(-sum) %>%
  ungroup( )

# case number
lag_10yrs %>%
dplyr::filter(ck == 1) %>%
  group_by(nr) %>%
  count() %>%
  nrow()

# control number
lag_10yrs %>%
dplyr::filter(ck == 0,
              lags == 1) %>%
  count()
  

class(lag_10yrs)


```


## 2. Parish SES

- I added SES to each of the datasets 
- Johnni sent us the Parish SES data which is percentage of population in various educational levels. We combined groups to estimate percentage of the population with a high school degree or higher

```{r}
# load parish SES
load(paste0(data.folder,'parish_ses/sogn_udd_2009b_1aR')) 

parish_ses <- sogn_udd_2009b_1aR %>%
  rowwise() %>% 
  mutate(parish_ses = sum(ses23,ses4,ses567,ses89, na.rm=TRUE)) %>% # create ses variable
  dplyr::select(-ses0, -ses1, -ses23, -ses4, -ses567, -ses89, -sesto) 
sogn_udd_2009b_1aR = NULL # remove parish SES r file

# load look-up for each participant's parish code

parish_lookup = haven::read_dta(paste0(data.folder,'parish_ses/ALS_geo_adr_1989_2013_1989_5_2.dta'))

# merge parish code at index date for each participant, then merge parish ses info 
data_1yr = left_join(data_1yr, parish_lookup, by=c('nr','lobnr','ck'))
data_1yr = left_join(data_1yr, parish_ses, by=c('sognecode'))
summary(data_1yr)

write_csv(data_1yr, paste0(yanelli_datasets, "one_yr_avg_data_ses.csv")) 


data_5yrs = left_join(data_5yrs, parish_lookup, by=c('nr','lobnr','ck'))
data_5yrs = left_join(data_5yrs, parish_ses, by=c('sognecode'))
summary(data_5yrs)


write_csv(data_5yrs, paste0(yanelli_datasets, "five_yr_avg_data_ses.csv")) 

data_10yrs = left_join(data_10yrs, parish_lookup, by=c('nr','lobnr','ck'))
data_10yrs = left_join(data_10yrs, parish_ses, by=c('sognecode'))
summary(data_10yrs)

write_csv(data_10yrs, paste0(yanelli_datasets, "ten_yrs_avg_data_ses.csv")) 

lag_10yrs = left_join(lag_10yrs, parish_lookup, by=c('nr','lobnr','ck'))
lag_10yrs = left_join(lag_10yrs, parish_ses, by=c('sognecode'))
summary(lag_10yrs)

write_csv(lag_10yrs, paste0(yanelli_datasets, "lagged_10yrs_parishSES.csv")) 

```

